---
title: "DDS_CaseStudy_2"
output:
  html_document:
    toc: true
    toc_depth: 3
author: "Kevin Boyd & Shikha Pandey"
date: "12/7/2021"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

DDSAnalytics is an analytics company that specializes in talent management solutions for Fortune 100 companies. It is planning to leverage data science for talent management to gain a competitive edge over its competition. Their executive leadership has identified predicting employee turnover as its first application of data science for talent management and they have tasked our data science team to conduct an analysis of existing employee data.

We have to use R on a given dataset to do a data analysis to identify factors that lead to attrition. Based on statistical analysis and evidence, we will identify the top three factors that contribute to turnover along with any other interesting trends and observations. We will also build models to predict attrition and salary.

# Video
[Attrition Case Study](https://youtu.be/MuLEzt19qhw)

# Load Libraries and Set Color Scheme
```{r message=F, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(GGally)
library(cowplot)
library(class)
library(caret)
library(e1071)
library(reshape2)
library(stringr)
library(naniar)
library(skimr) #for data summary
Rcolors <- c("lightseagreen", "darkslateblue","orange")
```


# Load and look at the data
```{r}
# The dataset has 870 observations (rows) and 36 variables (columns).
employee.df <- read.csv("CaseStudy2-data.csv", header = T)
str(employee.df)
skim(employee.df)
# There are 300 observations in each dataset below to predict Attrition and Salary respectively.
NoAttrition = read.csv("CaseStudy2CompSet No Attrition.csv", header = TRUE)
NoSalary <- read.csv("CaseStudy2CompSet(NoSalary).csv", header = T)
```


# Employee EDA
## Look at the Variables
### Investigate Missing Values in Each Column
```{r}
# No missing values are found for any variables (columns) in the dataset.
gg_miss_var(employee.df, show_pct = T) + labs(title = 'Percent Missing Values') + theme(plot.title = element_text(hjust = 0.5))
```

### Attrition Distribution
```{r}
employee.df %>% ggplot(aes(x = Attrition, fill = Attrition)) + 
  geom_bar() + 
  ggtitle("Distribution of Attrition")  + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_manual(values = Rcolors)
```

### Monthly Income
```{r}
# We see from the histogram that most number of employees have salary in range $2000-$4000 and then second highest range is $4000-$6000.
ggplot(employee.df, aes(x=MonthlyIncome, fill = Attrition)) +
  geom_histogram() +
  ggtitle("Monthly Income Based on Attrition") +
  xlab("Monthly Income") +
  ylab("Count") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_manual(values = Rcolors)
```

### Monthly Income by Groups
```{r}
# Let's create income groups and look at Attrition for those income groups.
employee.df$IncomeGroup <- cut(employee.df$MonthlyIncome, breaks = c(0, 2000, 4000, 6000, 10000, 15000, 20000), labels = c("<$2000","$2000-$4000","$4000-$6000","$6000-$10000","$10000-$15000","$15000-$20000"))

# Lowest Income Group (<$2000) has the highest Attrition rate, followed by second lowest Income Group ($2000-$4000).
employee.df %>% ggplot(aes_string(x = "IncomeGroup", fill = "Attrition")) + 
  geom_bar(position = "fill") + 
  scale_y_continuous(labels = scales::percent) + 
  labs(x = "Income Group", y = "Percent Employees", title = "Attrition for Income Groups") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_manual(values = Rcolors)
```

### Age by Groups
```{r}
# Let's create age groups and look at Attrition for those age groups.
employee.df$AgeGroup <- cut(employee.df$Age, breaks = c(18,25,35,45,60), labels = c("18-25","25-35","35-45","45-60"), include.lowest = TRUE)

# Younger age groups 18-25 and 25-35 have higher attrition rate.
employee.df %>% ggplot(aes_string(x = "AgeGroup", fill = "Attrition")) + 
  geom_bar(position = "fill") + 
  scale_y_continuous(labels = scales::percent) + 
  labs(x = "Age Group", y = "Percent Employees", title = "Attrition for Age Groups") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_manual(values = Rcolors)
```

### Job Role
```{r}
# Research Directors have highest attrition rate whereas Sales Representatives have the least. 
employee.df %>% ggplot(aes_string(x = "JobRole", fill = "Attrition")) + 
  geom_bar(position = "fill") + 
  scale_y_continuous(labels = scales::percent) + 
  labs(x = "Job Role", y = "Percent Employees", title = "Attrition for Job Role") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip() + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 12)) + 
  scale_fill_manual(values = Rcolors)
```

### Monthly Income based on Job Role
```{r}
ggplot(employee.df, aes(x=MonthlyIncome, fill = Attrition)) +
  geom_histogram() +
  facet_wrap(~JobRole) +
  ggtitle("Monthly Income based on Job Role") +
  xlab("Monthly Income") +
  ylab("Count") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_manual(values = Rcolors)
```


### Job Role and Job Satisfaction
```{r}
ggplot(employee.df, aes(x=JobSatisfaction, fill = Attrition)) +
  geom_bar() +
  ggtitle("Attrition based on Job Role and Job Satisfaction") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_manual(values = Rcolors) +
  facet_wrap(~JobRole)
```


### Job Role, Age, and Monthly Income 
```{r}
ggplot(employee.df, aes(x=Age, y=MonthlyIncome, color = Attrition)) +
  geom_point() +
  facet_wrap(~JobRole) +
  ggtitle("Attrition based on Age and Monthly Income") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = Rcolors)
```


### Overtime
```{r}
# Higher attrition rate for employees who work overtime.
employee.df %>% ggplot(aes_string(x = "OverTime", fill = "Attrition")) + 
  geom_bar(position = "fill") + 
  scale_y_continuous(labels = scales::percent) + 
  labs(x = "Overtime", y = "Percent Employees", title = "Attrition for Overtime") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_fill_manual(values = Rcolors)
```


### Business Travel
```{r}
# Attrition level goes up from Non-Travel to Travel-Frequently category.
employee.df %>% ggplot(aes_string(x = "BusinessTravel", fill = "Attrition")) + 
  geom_bar(position = "fill") + 
  scale_y_continuous(labels = scales::percent) + 
  labs(x = "Business Travel", y = "Percent Employees", title = "Attrition for Business Travel") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_fill_manual(values = Rcolors)
```


### Marital Status
```{r}
# Divorce contribute to attrition
bar <- ggplot(employee.df, aes(x=MaritalStatus, fill = MaritalStatus)) +
  geom_bar() +
  facet_wrap(~Attrition) +
  ggtitle("Attrition Dependent on Marital Status ") +
  xlab("Attrition") +
  ylab("Count")
bar <- bar + scale_fill_manual(values = Rcolors)

# Yes Attrition
# Make new data frames for married, divorced, and single
df1 <- employee.df %>%
  filter(Attrition == "Yes")
df1length <- NROW(df1)
df1married <- length(grep("Married", df1$MaritalStatus))
df1divorced <- length(grep("Divorced", df1$MaritalStatus))
df1single <- length(grep("Single", df1$MaritalStatus))

Yes_df <- data.frame(MaritalStatus = c("Married","Divorced","Single"), 
                     Value = c(df1married,df1divorced,df1single))

plot1 <- ggplot(Yes_df, aes(x="", y=Value, fill=MaritalStatus)) +
  geom_bar(width = 1, stat = "identity") + xlab("") + ylab("") + ggtitle("Attrition: Yes")
pie1 <- plot1 + coord_polar("y", start=0)
pie1 <- pie1 + scale_fill_manual(values = Rcolors)

# No Attrition
# Make new data frames for married, divorced, and single
df2 <- employee.df %>%
  filter(Attrition == "No")
df2length <- NROW(df2)
df2married <- length(grep("Married", df2$MaritalStatus))
df2divorced <- length(grep("Divorced", df2$MaritalStatus))
df2single <- length(grep("Single", df2$MaritalStatus))

No_df <- data.frame(MaritalStatus = c("Married","Divorced","Single"), 
                     Value = c(df2married,df2divorced,df2single))

plot2 <- ggplot(No_df, aes(x="", y=Value, fill=MaritalStatus)) +
  geom_bar(width = 1, stat = "identity") + xlab("") + ylab("") + ggtitle("Attrition: No")
pie2 <- plot2 + coord_polar("y", start=0)
pie2 <- pie2 + scale_fill_manual(values = Rcolors)

piecharts <- plot_grid(pie2,pie1, ncol = 2,  labels = c("B","C"))
allplots <- plot_grid(bar,piecharts, nrow = 2, labels = "A")
allplots
```


### Bar Plots Multiple Variables
```{r}
EDA1 <- ggplot(employee.df, aes(x=RelationshipSatisfaction, fill = Attrition)) +
  geom_bar(position = "fill") +
  ggtitle("Relationship Satisfaction") +
  ylab("% Count") + 
  scale_y_continuous(labels = scales::percent) + 
  theme(legend.position = "none") +
  scale_fill_manual(values = Rcolors)
EDA1 <- EDA1 + theme(plot.title = element_text(size = 12, hjust = 0.5))

EDA2 <- ggplot(employee.df, aes(x=JobLevel, fill = Attrition)) +
  geom_bar(position = "fill") +
  ggtitle("Job Level") +
  ylab("% Count") + 
  scale_y_continuous(labels = scales::percent) + 
  theme(legend.position = "none") +
  scale_fill_manual(values = Rcolors)
EDA2 <- EDA2 + theme(plot.title = element_text(size = 12, hjust = 0.5))

EDA3 <- ggplot(employee.df, aes(x=JobSatisfaction, fill = Attrition)) +
  geom_bar(position = "fill") +
  ggtitle("Job Satisfaction") +
  ylab("% Count") + 
  scale_y_continuous(labels = scales::percent) + 
  theme(legend.position = "none") +
  scale_fill_manual(values = Rcolors)
EDA3 <- EDA3 + theme(plot.title = element_text(size = 12, hjust = 0.5))

EDA4 <- ggplot(employee.df, aes(x=YearsWithCurrManager, fill = Attrition)) +
  geom_bar(position = "fill") +
  ggtitle("Years With Current Manager") +
  ylab("% Count") + 
  scale_y_continuous(labels = scales::percent) + 
  theme(legend.position = "none") +
  scale_fill_manual(values = Rcolors)
EDA4 <- EDA4 + theme(plot.title = element_text(size = 12, hjust = 0.5))

EDA5 <- ggplot(employee.df, aes(x=YearsSinceLastPromotion, fill = Attrition)) +
  geom_bar(position = "fill") +
  ggtitle("Years Since Last Promotion") +
  ylab("% Count") + 
  scale_y_continuous(labels = scales::percent) + 
  theme(legend.position = "none") +
  scale_fill_manual(values = Rcolors)
EDA5 <- EDA5 + theme(plot.title = element_text(size = 12, hjust = 0.5))

EDA6 <- ggplot(employee.df, aes(x=YearsAtCompany, fill = Attrition)) +
  geom_bar(position = "fill") +
  ggtitle("Years At Company") +
  ylab("% Count") + 
  scale_y_continuous(labels = scales::percent) + 
  theme(legend.position = "none") +
  scale_fill_manual(values = Rcolors)
EDA6 <- EDA6 + theme(plot.title = element_text(size = 12, hjust = 0.5))

EDA7 <- ggplot(employee.df, aes(x=YearsInCurrentRole, fill = Attrition)) +
  geom_bar(position = "fill") +
  ggtitle("Years in Current Role") +
  ylab("% Count") + 
  scale_y_continuous(labels = scales::percent) + 
  theme(legend.position = "none") +
  scale_fill_manual(values = Rcolors)
EDA7 <- EDA7 + theme(plot.title = element_text(size = 12, hjust = 0.5))

EDA8 <- ggplot(employee.df, aes(x=NumCompaniesWorked, fill = Attrition)) +
  geom_bar(position = "fill") +
  ggtitle("Companies Worked") +
  ylab("% Count") + 
  scale_y_continuous(labels = scales::percent) + 
  theme(legend.position = "none") +
  scale_fill_manual(values = Rcolors)
EDA8 <- EDA8 + theme(plot.title = element_text(size = 12, hjust = 0.5))

EDA9 <- ggplot(employee.df, aes(x=MaritalStatus, fill = Attrition)) +
  geom_bar(position = "fill") +
  ggtitle("Marital Status") +
  ylab("% Count") + 
  scale_y_continuous(labels = scales::percent) + 
  theme(legend.position = "none") +
  scale_fill_manual(values = Rcolors)
EDA9 <- EDA9 + theme(plot.title = element_text(size = 12, hjust = 0.5))

EDA10 <- plot_grid(EDA1,EDA2,EDA3,EDA4,EDA5,EDA6,EDA7,EDA8,EDA9, ncol = 3, nrow = 3)
EDA10
```


### Scatter Plots with Continuous Variables
```{r}
EDA.A <- ggplot(employee.df, aes(x=TotalWorkingYears, y=PercentSalaryHike, color = Attrition)) +
  geom_point() +
  ggtitle("Total Working Years vs % Salary Hike") +
  scale_color_manual(values = Rcolors) +
  ylab("% Salary Hike") + 
  theme(plot.title = element_text(size = 12, hjust = 0.5))

EDA.B <- ggplot(employee.df, aes(x=YearsWithCurrManager, y=PercentSalaryHike, color = Attrition)) +
  geom_point() +
  ggtitle("Yrs With Manager vs % Salary Hike") +
  scale_color_manual(values = Rcolors) +
  ylab("% Salary Hike") +
  xlab("Yrs With Manager") + 
  theme(plot.title = element_text(size = 12, hjust = 0.5))

EDA.C <-ggplot(employee.df, aes(x=YearsAtCompany, y=YearsInCurrentRole, color = Attrition)) +
  geom_point() +
  ggtitle("Yrs At Company vs Yrs In Role") +
  scale_color_manual(values = Rcolors) +
  ylab("Yrs In Role") + 
  theme(plot.title = element_text(size = 12, hjust = 0.5))

EDA.E <- ggplot(employee.df, aes(x=Age, y=YearsSinceLastPromotion, color = Attrition)) +
  geom_point() +
  ggtitle("Age vs Job Yrs Last Promotion") + 
  scale_color_manual(values = Rcolors) +
  ylab("Yrs Last Promotion") + 
  theme(plot.title = element_text(size = 12, hjust = 0.5))

EDA.G <- plot_grid(EDA.A,EDA.B,EDA.C,EDA.E, ncol = 2, nrow = 2)
EDA.G
```


# Models
## Attrition Models
### Naive Bayes
```{r}
set.seed(9)
NB.df <- select(employee.df, "Attrition", "Age", "Department", "EnvironmentSatisfaction", "JobInvolvement", "JobLevel", "JobRole", "JobSatisfaction", "MaritalStatus", "MonthlyIncome", "NumCompaniesWorked", "OverTime", "StockOptionLevel", "TotalWorkingYears", "WorkLifeBalance", "YearsAtCompany", "YearsInCurrentRole", "YearsWithCurrManager")
NBsplitPerc <- .85
NBtrainIndices <- sample(1:dim(NB.df)[1],round(NBsplitPerc * dim(NB.df)[1]))
NBdfTrain <- NB.df[NBtrainIndices,]
NBdfTest <- NB.df[-NBtrainIndices,]

# Naive Bayes model
NBmodel = naiveBayes(Attrition~.,data = NBdfTrain)
confusionMatrix(table(predict(NBmodel,NBdfTest),NBdfTest$Attrition))

# Make predictions with model
NBpreds = predict(NBmodel, NoAttrition)
NBpreds = data.frame(NBpreds)
NoAttrition$Attrition = NBpreds$NBpreds

Attrition.Classify = NoAttrition %>% select("ID", "Attrition") %>% arrange(ID)

# This code was used to label the predicted attrition on the file provided, but commented out to not replace the file when knitted at a later time.
#write.csv(Attrition.Classify, file = "Case2PredictionsBoyd_Pandey Attrition.csv", row.names=FALSE)

# plot counts of attrition
ggplot(Attrition.Classify, aes(x=Attrition, fill=Attrition)) +
  geom_bar(fill = Rcolors[c(1,2)]) +
  ggtitle("Predicted Attrition") +
  ylab("Count") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Multiple Train/Test Sets
```{r}
iterations = 100

masterAcc = matrix(nrow = iterations)
masterSens = matrix(nrow = iterations)
masterSpec = matrix(nrow = iterations)

splitPerc = .85 #Training / Test split Percentage

for(j in 1:iterations)
{
  
  trainIndices = sample(1:dim(NB.df)[1],round(splitPerc * dim(NB.df)[1]))
  train = NB.df[trainIndices,]
  test = NB.df[-trainIndices,]
  
  model = naiveBayes(train[,c(2:18)],train$Attrition)
  table(predict(model,test[,c(2:18)]),test$Attrition)
  CM = confusionMatrix(table(predict(model,test[,c(2:18)]),test$Attrition))
  masterAcc[j] = CM$overall[1]
  masterSens[j] = CM$byClass[1]
  masterSpec[j] = CM$byClass[2]
}

# Accuracy measurements
MeanAcc = colMeans(masterAcc)
MeanAcc
masterAcc <- data.frame(masterAcc)
masterAcc$TestNumber <- 1:nrow(masterAcc)
max(masterAcc$masterAcc)
min(masterAcc$masterAcc)
ggplot(masterAcc, aes(x=seq(1,100,1), y=masterAcc)) + 
  geom_line(color="darkslateblue") +
  ggtitle("Accuracy of 100 Train/Test Sets") +
  xlab("Number of Train/Test Sets") +
  ylab("Accuracy") + 
  theme(plot.title = element_text(hjust = 0.5))

# Sensitivity measurements
MeanSens = colMeans(masterSens)
MeanSens
masterSens <- data.frame(masterSens)
masterSens$TestNumber <- 1:nrow(masterSens)
max(masterSens$masterSens)
min(masterSens$masterSens)
ggplot(masterSens, aes(x=seq(1,100,1), y=masterSens)) + 
  geom_line(color="darkslateblue") +
  ggtitle("Sensitivity of 100 Train/Test Sets") +
  xlab("Number of Train/Test Sets") +
  ylab("Sensitivity") + 
  theme(plot.title = element_text(hjust = 0.5))

# Specificity measurements
MeanSpec = colMeans(masterSpec)
MeanSpec
masterSpec <- data.frame(masterSpec)
masterSpec$TestNumber <- 1:nrow(masterSpec)
max(masterSpec$masterSpec)
min(masterSpec$masterSpec)
ggplot(masterSpec, aes(x=seq(1,100,1), y=masterSpec)) + 
  geom_line(color="darkslateblue") +
  ggtitle("Specificity of 100 Train/Test Sets") +
  xlab("Number of Train/Test Sets") +
  ylab("Specificity") + 
  theme(plot.title = element_text(hjust = 0.5))

# Plot together
masterAcc$Value <- masterAcc$masterAcc
masterAcc$Test <- "Accuracy"
masterSpec$Value <- masterSpec$masterSpec
masterSpec$Test <- "Specificity"
masterSens$Value <- masterSens$masterSens
masterSens$Test <- "Sensitivity"
NB.df <- rbind(masterAcc[,c(2,3,4)],masterSens[,c(2,3,4)],masterSpec[,c(2,3,4)])


ggplot(NB.df, aes(x=TestNumber, y=Value, linetype=Test, colors=Test)) + 
  geom_line(color = "darkslateblue") +
  ggtitle("Accuracy, Sensitivity, & Specificity of 100 Train/Test Sets") +
  xlab("Number of Train/Test Sets") +
  ylab("Statistical Values") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = Rcolors)
```


### k-NN Models
```{r}
kNNsplitPerc <- .7

# Create dataframe for k-NN model 
knn.df <- employee.df

# Create standardized variables
knn.df$z_Age = scale(knn.df$Age)
knn.df$z_YearsSinceLastPromotion = scale(knn.df$YearsSinceLastPromotion)

kNNtrainIndices <- sample(1:dim(knn.df)[1],round(kNNsplitPerc * dim(knn.df)[1]))
kNNdfTrain <- knn.df[kNNtrainIndices,]
kNNdfTest <- knn.df[-kNNtrainIndices,]

# Internal Model
classifications <- knn.cv(knn.df[,c(2,25)],knn.df$Attrition, k = 20)
confusionMatrix(table(classifications,knn.df$Attrition))

# External Model
classifications1 <- knn(kNNdfTrain[,c(2,35)], kNNdfTest[,c(2,35)], kNNdfTrain$Attrition, 
                       prob = TRUE, k = 20)

confusionMatrix(table(classifications1,kNNdfTest$Attrition))
```


## Salary Model
### Linear Regression
```{r}
# Train and Test for Linear Regression
set.seed(9)
lmsplitPerc <- .8
trainIndices <- sample(1:dim(employee.df)[1],round(lmsplitPerc * dim(employee.df)[1]))
dfTrain <- employee.df[trainIndices,]
dfTest <- employee.df[-trainIndices,]
dfTrain <- na.omit(dfTrain)
dfTest <- na.omit(dfTest)

# Multiple variable linear regression with train and test
Model1_fit = lm(MonthlyIncome~Age+JobLevel+JobRole, data = dfTrain)
summary(Model1_fit)

# Show residuals
hist(Model1_fit$residuals, col = "darkslateblue", main = "Histogram of Residuals")
sqrt(sum((Model1_fit$residuals)^2))

# Predict with this model
preds2 <- predict(Model1_fit, newdata = dfTest)

# Assess model
MSPE = data.frame(Observed = dfTest$MonthlyIncome, Predicted = preds2)
MSPE$Resisdual = MSPE$Observed - MSPE$Predicted
MSPE$SquaredResidual = MSPE$Resisdual^2
mean(MSPE$SquaredResidual)

# Multiple variable linear regression
Model2_fit = lm(MonthlyIncome~Age+JobLevel+JobRole, data = employee.df)
summary(Model2_fit)

# Show residuals
hist(Model2_fit$residuals, col = "darkslateblue", main = "Histogram of Residuals")
sqrt(sum((Model2_fit$residuals)^2))

# Predict with this model
preds2 <- predict(Model2_fit, newdata = NoSalary)
preds2 <- as.data.frame(preds2)
NoSalary$Salary <- preds2[,1]

# Plot predicted values
NoSalary %>% ggplot(aes(x = Age, y = Salary)) + 
  geom_point(color = "darkslateblue") + 
  ggtitle("Predicted Salary by Age") +
  ylab("Salary") + 
  theme(plot.title = element_text(hjust = 0.5))

#Salary.Classify = NoSalary %>% select("ID", "Salary") %>% arrange(ID)

# This code was used to label the predicted Salary amounts on the file provided, but commented out to not replace the file when knitted at a later time.
#write.csv(Salary.Classify, file = "Case2PredictionsBoyd_Pandey Salary.csv", row.names=FALSE)
```


# Conclusion
1. Top 3 factors that contribute to employee turnover are Age, Monthly Income, and Job Role.
2. Sales Representative have the highest turnover.
3. Manufacturing Director, Research Director and Managers are the oldest, have highest income and have the least amount of turnover.
4. We used a Naïve Bayes model to predict attrition with 87% Accuracy.
5. We used a Linear Regression model to predict salary with RMSE of 1080 and extremely low P-Value (< 2.2e-16).
