---
title: 'Case Study 1: Beers and Breweries'
author: "Shikha Pandey"
date: "October 6, 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries and setting the theme for plots
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(stringi)
library(stringr)
library(naniar)
library(maps)
library(mapproj)
library(mice) #mice
library(class) #knn
library(caret) #confusionMatrix
library(e1071) #Naive Bayes
theme_set(theme_classic()) #Set the theme for plots
```

## Read in both Beer and Brewery files related to this case study.
```{r}
#Read in data from Beers.csv
Beers = read.csv(file.choose(),header = TRUE, na.strings = c(""," ","NA"))
#Read in data from Breweries.csv
Breweries = read.csv(file.choose(),header = TRUE, na.strings = c(""," ","NA"))
Breweries$State = as.factor(str_remove_all(Breweries$State, "\\s")) #Remove space preceeding State
```
```{r}
dim(Beers)
#Beer data has 2410 observations and 7 columns:

#1. Name: Name of the beer
#2. Beer_ID: Unique identifier of the beer
#3. ABV: Alcohol by volume of the beer
#4. IBU: International Bitterness Units of the beer
#5. Brewery_id: Brewery identifier associated with the beer
#6. Style: Style of the beer
#7. Ounces: Ounces of the beer
```
```{r}
dim(Breweries)
#Breweries data has 558 observations and 4 columns:

#1. Brew_ID: Unique identifier of the brewery
#2. Name: Name of the brewery
#3. City: City where brewery is located
#4. State: U.S. State where brewery is located
```

##1. Number of breweries present in each state.
```{r}
#Number of breweries by state
Breweries %>% group_by(State) %>% summarize(Count = n()) %>% arrange(desc(Count)) %>% print(n=Inf)

#Plot the map of all States in dataset with gradient color based on count of breweries
lookup = data.frame(State_Cd = state.abb, State = state.name) #makes a data frame with State name and state code
colnames(Breweries)[4] = "State_Cd" # Change Column Name
Breweries2 = merge(Breweries,lookup,"State_Cd") #make one dataset with state names and abb
BreweriesMapData = count(Breweries2,State) #count up the occurance of each state. 
colnames(BreweriesMapData)[2] = "Count" #change "n" to "Count"
BreweriesMapData$region <- tolower(BreweriesMapData$State)
BreweriesMapData2 = BreweriesMapData[-1]
states <- map_data("state")
map.df <- merge(states,BreweriesMapData2, by="region", all.x=T)
map.df <- map.df[order(map.df$order),]
Num_pos <- data.frame(region=tolower(state.name), long=state.center$x, lat=state.center$y)
Num_pos <- merge(Num_pos, BreweriesMapData2, by="region", all.x = TRUE)
#Plot
ggplot(map.df, aes(x=long,y=lat))+
  geom_polygon(aes(group=group, fill=Count),colour="white")+
  geom_text(data=Num_pos, aes(long, lat, label=Count)) +
  scale_fill_gradient(low = "lightsteelblue1",
                      high = "Orange",
                      guide = "colourbar")+
  ggtitle("Number of Breweries in each State") +
  coord_map()
#There are total 558 breweries in the US. CO has the most with 47 breweries. DC, ND, SD and WV are at the bottom with 1 brewery each.
```

##2.	Merge beer data with the breweries data. First 6 observations and the last 6 observations printed to check the merged file.
```{r}
#Merge the data frames for Beers and Breweries
Budweiser = merge(Beers,Breweries, by.x = "Brewery_id", by.y = "Brew_ID")
#Assigning new names to the columns of the data frame
colnames(Budweiser) <- c('Brewery_ID','Beer','Beer_ID', 'ABV', 'IBU', 'Style', 'Ounces', 'Brewery', 'City', 'State')
#Rearranging columns in data frame
Budweiser = Budweiser %>% select(Brewery_ID, Brewery, Beer_ID, Beer, ABV, IBU, Style, Ounces, City, State)
str(Budweiser)
#First 6 observations of merged dataset
head(Budweiser)
#Last 6 observations of merged dataset
tail(Budweiser)
#After merging the beers and breweries datasets based on Brewery_ID from beers dataset and Brew_ID from breweries dataset, resulting dataset, Budweiser, now has 2410 observations with 10 columns.
#Head of the dataset displays first 6 rows of data and tail of dataset displays last 6 rows of data from this merged dataset.
```

##3. Addressing the missing values in each column.
```{r}
#Looking for missing values in dataset
gg_miss_var(Budweiser) + labs(title = 'Missing Values', x = 'Dataset Columns')
sapply(Budweiser, function(x) sum(is.na(x)))
#Missing values are found for IBU (1005), ABV (62) and Style (5). It is a significant number for IBU (41.7%), notable for ABV (2.57%) but insignificant for Style (0.2%), so, I decided to exclude the missing 'Style' data (5 rows) from Budweiser.
```
```{r}
#Excluding missing data for 'Style'
Budweiser = Budweiser %>% filter(!is.na(Style))
sapply(Budweiser, function(x) sum(is.na(x)))
#After excluding 5 rows of missing 'Style' data, I notice that missing IBU and ABV are 1002 and 59 respectively as compared to 1005 and 62 earlier.
```
```{r}
#Assumption is that missing values of ABV and IBU are Missing At Random (MAR). Applying multiple imputation with function MICE() for ABV and IBU with covariant 'Style'.

#Multiple imputation for missing values of ABV and IBU
mice_mod <- mice(Budweiser[, c("IBU","Style")], method='pmm')
mice_complete1 <- complete(mice_mod)

mice_mod <- mice(Budweiser[, c("ABV","Style")], method='pmm')
mice_complete2 <- complete(mice_mod)

#Transfer the predicted missing values for ABV and IBU into the Budweiser dataset
Budweiser$IBU = mice_complete1$IBU
Budweiser$ABV = mice_complete2$ABV
```
```{r}
#Looking for missing values in dataset after imputation, none found
gg_miss_var(Budweiser) + labs(title = 'Missing Values', x = 'Dataset Columns')
sapply(Budweiser, function(x) sum(is.na(x)))
#With these imputed values, dataset is now complete with no more missing values.
```

##4. Median alcohol content and international bitterness unit for each state. Bar chart plots to compare.
```{r}
#Median acohol content and international bitterness unit for each state
Budweiser %>% group_by(State) %>% summarize(medianABV = median(ABV), medianIBU = median(IBU), Count = n()) %>% print(n=Inf)

##Plot a bar chart for median ABV for each State
#Arrange (order) data by descending median ABV
Budweiser_ABV = Budweiser %>% group_by(State) %>% summarize(medianABV = median(ABV), medianIBU = median(IBU), Count = n()) %>% arrange(desc(medianABV))
#Reorder level of State factor based on descending median ABV
Budweiser_ABV$State = factor(Budweiser_ABV$State, level = Budweiser_ABV$State[order(-Budweiser_ABV$medianABV)])

#column plot of ordered results
Budweiser_ABV %>% ggplot(aes(x = State, y = medianABV)) + geom_bar(stat="identity", width = 0.4, fill="tomato2") + labs(title = 'Median ABV by State', subtitle = 'Ordered highest to lowest', y = 'Median ABV') + theme(axis.text.x = element_text(angle=65, vjust=0.6))

##Plot a bar chart for median IBU for each State
#Arrange (order) data by descending median IBU
Budweiser_IBU = Budweiser %>% group_by(State) %>% summarize(medianABV = median(ABV), medianIBU = median(IBU), Count = n()) %>% arrange(desc(medianIBU))
#Reorder level of State factor based on descending median IBU
Budweiser_IBU$State = factor(Budweiser_IBU$State, level = Budweiser_IBU$State[order(-Budweiser_IBU$medianIBU)])

#column plot of ordered results
Budweiser_IBU %>% ggplot(aes(x = State, y = medianIBU)) + geom_bar(stat="identity", width = 0.4, fill="tomato2") + labs(title = 'Median IBU by State', subtitle = 'Ordered highest to lowest', y = 'Median IBU') + theme(axis.text.x = element_text(angle=65, vjust=0.6))
#Kentucky (KY) has the highest median alcohol content with 6.5% ABV and Utah (UT) has the lowest median alcohol content with 4% ABV.
#Delaware (DE) has the highest median IBU with 66 units and Kansas (KS) has the lowest median IBU with 20 units.
```

##5. State with the maximum alcoholic (ABV) beer and with the most bitter (IBU) beer.
```{r}
#State with maximum alcoholic (ABV) beer
Budweiser %>% select(State, ABV) %>% filter(ABV == max(Budweiser$ABV))

#State with most bitter (IBU) beer
Budweiser %>% select(State, IBU) %>% filter(IBU == max(Budweiser$IBU))

#Plot the states on US map
US_states <- map_data("state")
colorado = US_states %>% filter(region == "colorado")
Oregon = US_states %>% filter(region == "oregon")

ggplot() + 
  geom_polygon(data = US_states, aes(x = long, y = lat, group = group),colour="white", fill="grey40") +
  geom_polygon(data = colorado, aes(x=long, y=lat, group = group), fill = "orange") +
  geom_polygon(data = Oregon, aes(x=long, y=lat, group = group),fill = "lightsteelblue") +
  coord_quickmap()
#Colorado (CO) has the maximum alcoholic beer with 12.8% ABV.
#Oregon (OR) has the most bitter beer with 138 units.
```

##6. Summary statistics and distribution of the ABV.
```{r}
#Summary statistics for ABV
summary(Budweiser$ABV)
#The lowest ABV is 0.1%, the highest is 12.8%. The Mean is 5.98% and the Median value is 5.6%.
```
```{r}
#Plot to visualize the distribution
Budweiser %>% ggplot(aes(x = ABV)) + geom_histogram(bins = 50, fill = "orange") + labs(title = 'AVB Distribution', x = 'Alcohol by Volume (ABV)', y = "Count of Beers") + geom_vline(xintercept = median(Budweiser$ABV), color ="lightsteelblue", size = 2) + geom_vline(xintercept = mean(Budweiser$ABV), color = "black", linetype = "dotted", size = 1)
#Right skewedness for ABV indicates that more beers are produced with lesser alcohol by volume of the beer.
```
```{r}
#AVB Distribution by Container
Budweiser %>% ggplot(aes(x = as.factor(Ounces), y = ABV)) + geom_boxplot(fill = "orange") + labs(title = 'AVB Distribution by Container', x = 'Fluid Ounces', y = "Alcohol by Volume (ABV)")
#There are only maybe a handful of beers that come in 8.4 ounce and 16.9 ounce sizes and looks like they are strong beers. However, the 12 ounce containers have the lowest ABV median.
```

##7. Relationship between the bitterness of the beer and its alcoholic content shown by a scatter plot.
```{r}
#Scatter plot between ABV and IBU
Budweiser %>% ggplot(aes(x = ABV, y = IBU)) + geom_point() + geom_smooth(colour = "orange1") + labs(title = 'Relationship Between ABV and IBU', x = 'Alcohol by Volume (ABV)' , y = 'International Bitterness Unit (IBU)')
#Plot doesn't show an apparent relationship between ABV and IBU. It doesn't show that ABV increases for a higher IBU or vice versa.
```

##8. KNN classification to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale.
```{r}
#Create a dataset only for beers with 'Ale' or 'IPA' in their name from the Budweiser dataset
Budweiser_Ale = Budweiser %>% filter(str_detect(Style, "Ale") | str_detect(Style, "IPA"))

#Identify and label these beers as 'IPA' or 'Other Ales' in a new column
Budweiser_Ale = Budweiser_Ale %>% 
  mutate(Style_Ale = case_when(
    str_detect(Style, "India Pale Ale") | str_detect(Style, "IPA") ~ "IPA", 
               TRUE ~ "Other Ales")) %>%
  mutate(Style_Ale = factor(Style_Ale, levels = c("IPA", "Other Ales")))

#Plot indicating beer type with their ABV and IBU
Budweiser_Ale %>% ggplot(aes(x = ABV, y = IBU, color = Style_Ale)) + geom_point() + labs(title = 'Ale Type with IBU and ABV', color = 'Ale Type')
#IPAs and Other Ales form clusters in terms of their ABV and IBU values.
```

```{r}
#k-NN: Loop for many k and one training/test partition
set.seed(7)     #Set the seed
numks = 30      #Value for k (1 to 30)
splitPerc = .70 #Training/Test split Percentage

#Dataframe to store statistics - accuracy, sensitivity and specificity for each k value
Ale_Stat = data.frame(accuracy = numeric(30), sensitivity = numeric(30), specificity = numeric(30), positive_class = character(30), k = numeric(30))

trainIndices = sample(1:dim(Budweiser_Ale)[1],round(splitPerc * dim(Budweiser_Ale)[1]))
train = Budweiser_Ale[trainIndices,]
test = Budweiser_Ale[-trainIndices,] 

for(i in 1:numks)
{
  classifications = knn(train[,c(5,6)],test[,c(5,6)],train$Style_Ale,prob = TRUE, k = i)
  CM = confusionMatrix(table(classifications,test$Style_Ale))
  CM
  Ale_Stat$accuracy[i] = CM$overall[1]
  Ale_Stat$sensitivity[i] = CM$byClass[1]
  Ale_Stat$specificity[i] = CM$byClass[2]
  Ale_Stat$positive_class = CM$positive
  Ale_Stat$k[i] = i
}

#Reshape data frame from wide to long format
Ale_ggp <- data.frame(x = Ale_Stat$k,
                      y = c(Ale_Stat$accuracy, Ale_Stat$sensitivity, Ale_Stat$specificity),
                      Statistic = c(rep("Accuracy", nrow(Ale_Stat)),
                                rep("Sensitivity", nrow(Ale_Stat)),
                                rep("Specificity", nrow(Ale_Stat))))

#Plot the Statistics for k-value
ggplot(Ale_ggp, aes(x, y, col = Statistic)) + geom_line() + labs(title = 'Statistics vs k-value', x = 'k-value' , y = 'Statistical Value') + scale_x_continuous(breaks = seq(0,30,5)) + scale_y_continuous(breaks = seq(0,.95,.02)) + theme_grey()

#Find the k-value with maximum accuracy level
Ale_Stat[which.max(Ale_Stat$accuracy),]

#Mean Statistical values for k-values from 1 to 30
mean(Ale_Stat$accuracy)
mean(Ale_Stat$sensitivity)
mean(Ale_Stat$specificity)
#Based on ABV and IBU for the beers in this dataset, the Machine Learning k-NN Classification model classifies a beer with about 87% accuracy whether it's an IPA or other type of beer. The best k-value is 7 with highest accuracy of 89%, positive class as IPA.

  classifications = knn(train[,c(5,6)],test[,c(5,6)],train$Style_Ale,prob = TRUE, k = 15)
  CM = confusionMatrix(table(classifications,test$Style_Ale))
  CM
```

#Naive Bayes
```{r}
#naiveBayes()
model = naiveBayes(train[,c(5,6)],train$Style_Ale,laplace = 1)
CM = confusionMatrix(table(predict(model,test[,c(5,6)]),test$Style_Ale))
CM
#Another Machine Learning Naive Bayes Classification model compares with the k-NN with providing about 85% accuracy and positive classification as IPA.
```

#9.1 Insight - Which top 10 cities have the most number of breweries?
```{r}
#Create another column in Breweries dataset with City, Sate combination
Breweries$City_State = paste0(Breweries$City, ",", Breweries$State_Cd)

#Top 10 cities with most breweries
Breweries_Cty = Breweries %>% group_by(City_State) %>% summarize(Count = n()) %>% arrange(desc(Count)) %>% head(n=10)

#Reorder level of City, State combo factor based on descending brewery count
Breweries_Cty$City_State = factor(Breweries_Cty$City_State, level = Breweries_Cty$City_State [order(-Breweries_Cty$Count)])

##Plot a bar chart for City, State combo for each State
Breweries_Cty %>% ggplot(aes(x = City_State, y = Count)) + geom_bar(stat="identity", width = 0.5, fill="orange") + labs(title = 'Top 10 Cities with Most Breweries', x = 'City', y = 'Breweries') + scale_x_discrete(labels = function(x) str_wrap(x, width = 10))

#Portland, OR came up top in the United States with 11 breweries and San Francisco, CA gets the 10th spot with 5 breweries. Boulder, CO, Chicago, IL and Seattle, WA tie with 9 breweries each.
```

#9.2 Insight - Which top 10 breweries produce the most number of beers?
```{r}
#Create another column in Budweiser dataset with Brewery, City, Sate combination
Budweiser$Brew_Cty_State = paste0(Budweiser$Brewery," ",Budweiser$City,",", Budweiser$State)

#Top 10 breweries that produce the most number of beers
Budweiser_Brew = Budweiser %>% group_by(Brew_Cty_State) %>% summarize(Count = n()) %>% arrange(desc(Count)) %>% head(n=10)

#Reorder level of City, State combo factor based on descending brewery count
Budweiser_Brew$Brew_Cty_State = factor(Budweiser_Brew$Brew_Cty_State, level = Budweiser_Brew$Brew_Cty_State [order(-Budweiser_Brew$Count)])

##Plot a bar chart for City, State combo for each State
Budweiser_Brew %>% ggplot(aes(x = Brew_Cty_State, y = Count)) + geom_bar(stat="identity", width = 0.5, fill="orange") + labs(title = 'Top 10 Breweries with Most Beers', x = 'Brewery', y = 'Beers') + scale_x_discrete(labels = function(x) str_wrap(x, width = 10))

#Brewery Vivant in Grand Rapids, MI came up top producing 62 unique beers and Bonfire Brewing Company in Eagle, CO claimed 10th spot with producing 19 different beers.
```



